---
title: "m6Aboost Vignettes"
author:
- name: "You Zhou"
  affiliation: 
  - Buchmann Institute for Molecular Life Sciences, Frankfurt am Main, Germany
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
    BiocStyle::html_document:
        toc_float: true
    BiocStyle::pdf_document: default
package: cliProfiler
vignette: |
    %\VignetteIndexEntry{cliProfiler Vignettes}
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteEncoding{UTF-8}
---

```{r style, echo=FALSE, results='asis'}
BiocStyle::markdown()
```

# Introduction
N6-Methyladenosine (m6A) is the most abundant mRNA internal modification and 
associate with the many different aspects of mRNA's life, e.g. exporting, 
translation, stability and so on.   
  
In our paper, 
"[Deep and accurate detection of m6A RNA modifications using miCLIP2 and m6Aboost machine learning](https://www.biorxiv.org/content/10.1101/2020.12.20.423675v1)", we 
present a improved experiment protocol (miCLIP2) and a novel computational 
analysis pipeline (m6Aboost). This package includes a workflow and some 
functions for running the m6Aboost pipeline.

# m6Aboost Workflow
For running this workflow, the user may need to go through the process step in 
our pipeline to get the truncation read counts without the C to T transition 
and the the read counts of C to T transition. Then the user need to do the 
peakcalling step with 
[pureCLIP](https://github.com/skrakau/PureCLIP). After the peakcalling, user 
need to assign the truncation and the transition read counts to the peaks that 
output from `pureCLIP`. Please note that theoretically the m6A associated C to T 
transition would happens one nucleotide after the m6A sites, so when we assign 
the transitions we should assign the transition read to the peak one nucleotide 
before the transition place. Besides, the strand information should also be 
consider when we assign the transition read counts. These truncation and 
transition information should be assign to the peaks and store as the metadata 
in the `GRanges` object of peaks.  
  
Afterwards, user need to calculated the `mean value` of truncation and 
transition read counts among the replicates. 

## getFeatures
For training the m6Aboost, we used the sequence, transcript region and the 
miCLIP2 features. The miCLIP2 features includes relative signal strength (RSS) 
and C to T transition read counts. The function `getFeatures` could help user 
to get all these features that required by m6Aboost. Pleaes note the annotation 
file for running the function `getFeatures` should be downloaded from 
[GENCODE](https://www.gencodegenes.org/).

```{r initialize, results="hide", warning=FALSE, message=FALSE}
library(m6Aboost)
## For test all the functions we need to first load the test files
testpath <- system.file("extdata", package = "m6Aboost")
testpath2 <- system.file("extdata", package = "cliProfiler")
test_gff3 <- file.path(testpath2, "annotation_test.gff3")
test <- readRDS(file.path(testpath, "test.rds"))

test
```

```{r}
## get all the features for the m6Aboost
test<- getFeatures(test, test_gff3, "Reads_mean", "CtoT")
test
```

## m6Aboost
The function `m6Aboost` can be used to get the prediction result. For using 
this function, user just need to use the output object from `getFeatures` 
function as input and specify the name of `BSgenome` of the current species. 
Please note that the `BSgenome` which contain the sequence information should 
be downloaded in advance.

```{r}
out <- m6Aboost(test, "BSgenome.Mmusculus.UCSC.mm10")
out
```

The `class` in the meta data represents whether the classifier predict this 
site is a m6A sites or not. The `prob` means the prediction score of this site.










