---
title: "m6Aboost Vignettes"
author:
- name: "You Zhou"
  affiliation: 
  - Buchmann Institute for Molecular Life Sciences, Frankfurt am Main, Germany
- name: "Kathi Zarnack"
  affiliation: 
  - Buchmann Institute for Molecular Life Sciences, Frankfurt am Main, Germany
date: "`r format(Sys.time(), '%B %d, %Y')`"
bibliography: library.bib
output:
    BiocStyle::html_document:
        toc_float: true
    BiocStyle::pdf_document: default
package: m6Aboost
vignette: |
    %\VignetteIndexEntry{m6Aboost Vignettes}
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteEncoding{UTF-8}
---

```{r style, echo=FALSE, results='asis'}
BiocStyle::markdown()
```

# Introduction

N6-Methyladenosine (m6A) is the most abundant mRNA internal modification. It 
associates with the many different aspects of mRNA's life, e.g. exporting, 
translation, stability and so on.   
  
In our paper [@Koetel2020], 
we present a improved experiment protocol (**miCLIP2**) and a novel 
computational analysis pipeline (**m6Aboost**). This package includes a 
workflow and some functions for running the m6Aboost pipeline.

The `r Biocpkg("m6Aboost")` package is available at
[https://bioconductor.org](https://bioconductor.org) and can be
installed via `BiocManager::install`:

```{r BiocManager, eval=FALSE}
if (!require("BiocManager"))
    install.packages("BiocManager")
BiocManager::install("m6Aboost")
```

A package only needs to be installed once. Load the package into an
R session with

```{r initialize, results="hide", warning=FALSE, message=FALSE}
library(m6Aboost)
```

# m6Aboost Workflow

For running this workflow, the user may need to go through the process step in 
our pipeline to get the truncation read counts without the C to T transition 
reads. And the the read counts of C to T transitions. Then the user need to run 
the peakcalling step with [pureCLIP](https://github.com/skrakau/PureCLIP).  
  
## Loading test data set

For testing `getFeatures` and `m6Aboost` functions, we need 
to first load the test data set within the package. The expected input 
`GRanges` object for `getFeatures` should include the *mean value* of 
truncation read counts on the peak and the C to T transition read counts behind 
the peak. In the test data set, `Reads_mean` stands for the *mean value* of 
truncation read counts. `CtoT` means the *mean value* of C to T transition read 
counts. The test `test_gff3` is a subset of `gff3` file which download from 
[GENCODE](https://www.gencodegenes.org/).

```{r echo=TRUE, warning=FALSE, message=FALSE}
library(m6Aboost)
## For test all the functions we need to first load the test files
testpath <- system.file("extdata", package = "m6Aboost")
test_gff3 <- file.path(testpath, "test_annotation.gff3")
test <- readRDS(file.path(testpath, "test.rds"))

test
```

## Read counts assignment

After the peakcalling, user need to assign the truncation and the transition 
read count to the peaks that output from `pureCLIP`. In `r Biocpkg("m6Aboost")` 
we provide two functions `truncationAssignment` and `CtoTAssignment` to help 
user to assign the truncation and C to T transition read counts to the peaks.

```{r eval=FALSE, include=TRUE}
## truncationAssignment can help to assign the truncation read count
## Here show the usage of the assignment functions
## The input peak should be a GRanges object
peak <- truncationAssignment(peak, 
    bw_positive="/path/to/WT1_positive_truncation.bw",
    bw_negative="/path/to/WT1_negative_truncation.bw",
    sampleName = "WT1")

## CtoTAssignment can help to assign the truncation read count
peak <- CtoTAssignment(peak, 
    bw_positive="/path/to/WT1_positive_truncation.bw",
    bw_negative="/path/to/WT1_negative_truncation.bw",
    sampleName = "CtoT_WT1")
```

Afterwards, user need to calculated the *mean value* of truncation and 
transition read counts among the replicates. 

```{r eval=FALSE, include=TRUE}
## For example if the user have two replicates, user can easily calculate the 
## mean value like this way:
peak$WTmean <-(peak$WT1 + peak$WT2)/2
```

## Extract features for the m6Aboost model

For training the m6Aboost, we used the surranding sequence, transcript region 
and the miCLIP2 features. The miCLIP2 features includes relative signal 
strength (RSS) and C to T transition read counts. The function `getFeatures` 
could help user to get all these features that required by m6Aboost. Please 
note the function `getFeatures` require to use the annotation file that 
downloaded from [GENCODE](https://www.gencodegenes.org/).

```{r}
## Get all the features for the m6Aboost
test<- getFeatures(test, test_gff3, 
                   colname_reads="WTmean", 
                   colname_C2T="CtoTmean")
test
```

## Prediction

The function `m6Aboost` can be used to get the prediction result. For using 
this function, user just need to use the output object from `getFeatures` 
function as input and specify the name of `BSgenome` of the current species. 
Please note that the `BSgenome` which contain the sequence information should 
be downloaded from `Bioconductor` in advance.

```{r warning=FALSE, message=FALSE}
out <- m6Aboost(test, "BSgenome.Mmusculus.UCSC.mm10")
out
```

The *class* in the meta data represents whether the classifier predict this 
site is a m6A sites or not. The *prob* means the prediction score of this site. 

# Session info

```{r sessionInfo}
sessionInfo()
```

# References

